On the presentation of orthogonal groups over polynomial rings

### 2019-05-28 dl-paper.tex
Added couple of more lemmas

\documentclass[oneside, 8pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true,unicode]{hyperref}
\usepackage[hyperref=true, backend=biber, bibencoding=utf8, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage[notref,notcite]{showkeys}

@@

\subsection{The proof of the injectivity theorem}
Throughout this subsection $\Phi$ denotes arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise.
We denote by $B$ the ring $M[X] + A[X^{-1}]$ considered as a graded subring of $A[X, X^{-1}]$.
For shortness we also write $G_M^0$ instead of $\overline{\St}(\Phi, A, M)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]) &  } \] 
Denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of $\St(\Phi, A[X, X^{-1}]) \times \St(\Phi, A[X\inv]) \times (1 + M)^*$ consisting of those triples $(g, h, u)$ for which 
$p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}$ belongs to $\overline{G}_M^{\geq 0}$.
We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$.
It is clear that this action is well defined.
Denote by $\overline{V}$ the set of orbits of the action of $G_M^0$ on $V$.
Whenever $(g, h_1, u)$ and $(g, h_2, u)$ lie in the same orbit we use the notation $(g, h_1, u) \sim (g, h_2, u)$.

The next step of the proof is to define for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for those triples $(g, h, u)$ for which $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u)$ happens to belong to $P_\alpha(m)$ for some $m \in M$, then $t_\alpha(aX^{-1})$ is given by the following identity:
\[ t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\]

@@

\begin{lemma}\label{lem:orbit-action} Let $(A, M)$ be a local pair. 
Then for any $\alpha \in \Phi$ the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ gives rise to a well-defined total orbit function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $(g, h, u) \in V$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $t_\alpha(aX^{-1})$ is defined on $(g, h \cdot i_+(h_0), u)$, which is also a representative of the same orbit as $(g, h, u)$.

@@

The following proposition is a direct generalization of~\cite[Proposition~4.3]{Tu83}.
\begin{prop} The operators $T_\alpha(aX^{-1})$ satisfy all relations $\text{R1}_1, $ $\text{R2}_1$ and relations $\text{R3}_{1,1}$ in the special case when $\alpha, \beta$ form a sharp angle.
\end{prop}
\begin{proof}
Suppose $(g, h, u)$ is such that $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$.
Set $h' := x_\alpha\left(-\tfrac{aX^{-1}}{1+am}\right) \cdot h^{-1}_\alpha(1+am) \cdot x_\alpha\left(-bX^{-1}\left(1 + \tfrac{bm}{1+am}\right)^{-1}\right) =
  x_\alpha\left(- \tfrac{(a+b)X^{-1}}{1 + (a + b)m}\right) \cdot h^{-1}_\alpha(1+am).$
The assertion for $\text{R1}_1$ is obtained via~\cref{SR:additivity}:
\begin{multline} \nonumber T_\alpha(bX^{-1}) T_\alpha(aX^{-1}) \left(g,\ h,\ u\right) \sim \\ \sim
T_\alpha(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am),\ u(1 + am)\right) = \\
 = \left( x_\alpha\left((a+b)X^{-1}\right)\cdot g,\ h \cdot h',\ u(1 + am + bm) \right) \sim T_\alpha((a+b)X^{-1}) \left( g,\ h,\ u\right), \end{multline}

Let us verify $\text{R2}_1$. 
Let $\alpha, \beta$ be a pair of roots such that $\alpha + \beta \in \Phi$.
Notice that we could assume from the start that $p(g, h, u) \in j_+(K(A[X], M[X]) \cap P_\alpha(m) \cap P_{\alpha + \beta}(m'))$ for some $m,m' \in M$ (cf. the first part of the proof of~\cref{lem:orbit-action}).
Now for $a, b \in A$ set $\epsilon = N_{\alpha, \beta},$ $c = 1 - \epsilon abm'$.
Applying~\cref{SR:obtuse} to a pair of roots $\alpha, \alpha+\beta$ we obtain that 
\begin{multline} \nonumber
[T_\beta(b), T_\alpha(aX^{-1})] \left(g, h, u \right) = \\
 = T_\beta(b) T_\alpha(aX^{-1}) \left( x_\beta(-b) \cdot x_\alpha (-a X^{-1}) \cdot g, h \cdot x_\alpha\left(\tfrac{aX^{-1}}{1-am}\right), u (1 - am) \right) \sim \\
 \sim T_\beta(b) \cdot T_\alpha(aX^{-1}) \left( x_\beta(-b) \cdot x_\alpha (-a X^{-1}) g, h', u (1 - am) \right) = \\
 = \left([x_\beta(b), x_\alpha(aX^{-1})]\cdot g, h'', u c\right) \sim T_{\alpha+\beta}(-\epsilon abX^{-1})(g, h, u),\end{multline} 
 where $h' = h\cdot x_\alpha\left(\tfrac{a X^{-1}}{1-am}\right) \cdot h_{\alpha}^{-1}(1-am)\cdot h_{\beta}(c^{-1})\cdot x_{\beta}(bc)\cdot$ 
  and the last congruence is obtained as follows:
 \begin{multline} \nonumber
 h'' = h' \cdot x_\alpha\left(-\tfrac{a(1 - am)X^{-1}}{c}\right) \sim \\ \sim  h\cdot x_\alpha\left(\tfrac{a X^{-1}}{1-am}\right) \cdot h_{\alpha}^{-1}(1-am)\cdot h_{\beta}(c^{-1})\cdot x_{\alpha + \beta}\left(\epsilon ab(1-am)X^{-1}\right)\cdot x_\alpha\left(-\tfrac{a(1 - am)X^{-1}}{c}\right) = \\
 = h\cdot x_\alpha\left(\tfrac{a X^{-1}}{1-am}\right) \cdot x_{\alpha + \beta}\left(\tfrac{\epsilon \cdot abX^{-1}}{c}\right)\cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 - am}\right) \cdot h_{\alpha}^{-1}(1-am)\cdot h_{\beta}(c^{-1}) \sim
 h\cdot x_{\alpha + \beta}\left(\tfrac{\epsilon   abX^{-1}}{1 - \epsilon a b m'}\right).
\end{multline}

>>>>

\documentclass[oneside, 8pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true,unicode]{hyperref}
\usepackage[hyperref=true, backend=biber, bibencoding=utf8, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}
%\usepackage[notref,notcite]{showkeys}

@@

\subsection{The proof of the injectivity theorem}
Throughout this subsection $\Phi$ denotes arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise.
We denote by $B$ the ring $M[X] + A[X^{-1}]$ considered as a graded subring of $A[X, X^{-1}]$.
For shortness we also write $G_M^0$ instead of $\overline{\St}(\Phi, A, M)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]) &  } \] 
Denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of $\St(\Phi, A[X, X^{-1}]) \times \St(\Phi, A[X\inv]) \times (1 + M)^*$ consisting of those triples $(g, h, u)$ for which 
$p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}$ belongs to $\overline{G}_M^{\geq 0}$.
We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$.
It is clear that this action is well defined.
Denote by $\overline{V}$ the set of orbits of the action of $G_M^0$ on $V$.
Whenever $(g, h_1, u)$ and $(g, h_2, u)$ lie in the same orbit we use the notation $(g, h_1, u) \sim (g, h_2, u)$ and say that these triples are congruent.

The next step of the proof is to define for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for the triples $(g, h, u)$ satisfying $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u)$ happens to belong to $P_\alpha(m)$ for some $m \in M$, then $t_\alpha(aX^{-1})$ is given by the following identity:
\[ t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\]

@@

\begin{lemma}\label{lem:orbit-action} Let $(A, M)$ be a local pair. 
Then for any $\alpha \in \Phi$ and $a \in A$ the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ gives rise to a well-defined (total) function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $(g, h, u) \in V$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $t_\alpha(aX^{-1})$ is defined on $(g, h \cdot i_+(h_0), u)$, which is also a representative of the same orbit as $(g, h, u)$.

@@

For $a + Xf \in A + XM[X]$ we can define the operator $T_\alpha(a + Xf) \colon \overline{V} \to \overline{V}$ by the following identity:
\[ T_\alpha(a + Xf) \cdot (g, h, u) = (x_\alpha(a + Xf) \cdot g, h\cdot x_{\alpha}(-a), u).  \]
It is easy to see that this definition is correct.

\begin{lemma}
 The operators $T_\alpha$ satisfy $\text{R3}_{1,0}$.
\end{lemma}
\begin{proof}
 We have to show that $[T_\alpha(aX^{-1}), T_\beta(b)](g, h, u) = (g, h, u)$ for $\beta\in Z(\alpha)$, $a, b\in A$.
 \begin{multline}
  T_\beta(b) T_\alpha(aX^{-1}) T_\beta(-b) T_\alpha(-aX^{-1})(g, h, u) = \\
  = T_\beta(b) T_\alpha(aX^{-1}) \left(x_\beta(-b) x_\alpha(-aX^{-1}) \cdot g, h \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1-am}\right) \cdot x_\beta(b), u(1-am)\right)
 \end{multline}
\end{proof}

\begin{lemma}
 For $a, b\in A$, $f, g \in M[X]$, $\alpha \neq \pm \beta \in \Phi$ holds
 \[ T_\beta(f) T_\alpha(aX^{-1})  T_\beta(g) T_\alpha(bX^{-1})(g, h, u) \]
\end{lemma}
\begin{proof}
Suppose $(g, h, u)$ is such that $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$.
Set $h' := x_\alpha\left(-\tfrac{aX^{-1}}{1+am}\right) \cdot h^{-1}_\alpha(1+am) \cdot x_\alpha\left(-bX^{-1}\left(1 + \tfrac{bm}{1+am}\right)^{-1}\right) =
  x_\alpha\left(- \tfrac{(a+b)X^{-1}}{1 + (a + b)m}\right) \cdot h^{-1}_\alpha(1+am).$
The assertion for $\text{R1}_1$ is obtained via~\cref{SR:additivity}:
\begin{multline} \nonumber T_\alpha(bX^{-1}) T_\alpha(aX^{-1}) \left(g,\ h,\ u\right) \sim \\ \sim
T_\alpha(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am),\ u(1 + am)\right) = \\
 = \left( x_\alpha\left((a+b)X^{-1}\right)\cdot g,\ h \cdot h',\ u(1 + am + bm) \right) \sim T_\alpha((a+b)X^{-1}) \left( g,\ h,\ u\right), \end{multline}

\end{proof}

The following proposition is a direct generalization of~\cite[Proposition~4.3]{Tu83}.
\begin{prop} The operators $T_\alpha(aX^{-1})$ satisfy all relations $\text{R1}_1, $ $\text{R2}_{0,1}$ and relations $\text{R3}_{1,1}$ in the special case when $\alpha, \beta$ form a sharp angle.
\end{prop}
\begin{proof}
Suppose $(g, h, u)$ is such that $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$.
Set $h' := x_\alpha\left(-\tfrac{aX^{-1}}{1+am}\right) \cdot h^{-1}_\alpha(1+am) \cdot x_\alpha\left(-bX^{-1}\left(1 + \tfrac{bm}{1+am}\right)^{-1}\right) =
  x_\alpha\left(- \tfrac{(a+b)X^{-1}}{1 + (a + b)m}\right) \cdot h^{-1}_\alpha(1+am).$
The assertion for $\text{R1}_1$ is obtained via~\cref{SR:additivity}:
\begin{multline} \nonumber T_\alpha(bX^{-1}) T_\alpha(aX^{-1}) \left(g,\ h,\ u\right) \sim \\ \sim
T_\alpha(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am),\ u(1 + am)\right) = \\
 = \left( x_\alpha\left((a+b)X^{-1}\right)\cdot g,\ h \cdot h',\ u(1 + am + bm) \right) \sim T_\alpha((a+b)X^{-1}) \left( g,\ h,\ u\right), \end{multline}

Let us verify $\text{R2}_1$. 
Let $\alpha, \beta$ be a pair of roots such that $\alpha + \beta \in \Phi$.
Notice that we could assume from the start that $p(g, h, u) \in j_+(K(A[X], M[X]) \cap P_\alpha(m) \cap P_{\alpha + \beta}(m'))$ for some $m,m' \in M$ (cf. the first part of the proof of~\cref{lem:orbit-action}).
Now for $a, b \in A$ set $\epsilon = N_{\alpha, \beta},$ $c = 1 - \epsilon abm'$.
Applying~\cref{SR:obtuse} to the pair of roots $\alpha, \alpha+\beta$ we obtain that 
\begin{multline} \nonumber
[T_\beta(b), T_\alpha(aX^{-1})] \left(g, h, u \right) = \\
 = T_\beta(b) T_\alpha(aX^{-1}) \left( x_\beta(-b) \cdot x_\alpha (-a X^{-1}) \cdot g, h \cdot x_\alpha\left(\tfrac{aX^{-1}}{1-am}\right), u (1 - am) \right) \sim \\
 \sim T_\beta(b) \cdot T_\alpha(aX^{-1}) \left( x_\beta(-b) \cdot x_\alpha (-a X^{-1}) g, h', u (1 - am) \right) = \\
 = \left([x_\beta(b), x_\alpha(aX^{-1})]\cdot g, h'', u c\right) \sim T_{\alpha+\beta}(-\epsilon abX^{-1})(g, h, u),\end{multline} 
 where $h' = h\cdot x_\alpha\left(\tfrac{a X^{-1}}{1-am}\right) \cdot h_{\alpha}^{-1}(1-am)\cdot h_{\beta}(c^{-1})\cdot x_{\beta}(bc)\cdot$ 
  and the last congruence is obtained as follows:
 \begin{multline} \nonumber
 h'' = h' \cdot x_\alpha\left(-\tfrac{a(1 - am)X^{-1}}{c}\right) \sim \\ \sim  h\cdot x_\alpha\left(\tfrac{a X^{-1}}{1-am}\right) \cdot h_{\alpha}^{-1}(1-am)\cdot h_{\beta}(c^{-1})\cdot x_{\alpha + \beta}\left(\epsilon ab(1-am)X^{-1}\right)\cdot x_\alpha\left(-\tfrac{a(1 - am)X^{-1}}{c}\right) = \\
 = h\cdot x_\alpha\left(\tfrac{a X^{-1}}{1-am}\right) \cdot x_{\alpha + \beta}\left(\tfrac{\epsilon \cdot abX^{-1}}{c}\right)\cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 - am}\right) \cdot h_{\alpha}^{-1}(1-am)\cdot h_{\beta}(c^{-1}) \sim
 h\cdot x_{\alpha + \beta}\left(\tfrac{\epsilon   abX^{-1}}{1 - \epsilon a b m'}\right).
\end{multline}
